Lyman-$\alpha$ photons through rotating outflows

### 2018-08-08 paper/paper.tex
revised intro

Recent advances in instrumentation have revealed the presence of gas
rotation on vastly different physical scales.
For instance, spatially resolved spectra on compact dwarf galaxies
have measured clear signs of gas showing pure rotation kinematics
\citep{2015A&A...577A..21C,2017A&A...600A.125C} and the recent mapping of
high redshift circumgalactic regions have also revealed kinematic
evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
These systems with star formation, neutral gas and low dust
contents naturally produce a \lya emission line \citep{PartridgePeebles}
motivating the observational work to phenomenologically link
H$\alpha$ kinematics to \lya spectra \citep[e.g.][]{Herenz2016}.

@@

However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies are
more complex than pure rotation.
The case for outflows as traced in the \lya line is solidly
established.  
In many observations the \lya line profile has a single peak
redwards from the line's center, in other cases there is a double peak
but the peak on the red side is stronger
\citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
These features have been explained as the consequence of multiple
\lya photon scatterings through a homogeneous outflowing shell of
neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2015ApJ...812..123G}.

Nevertheless, a study of the combined effects of outflows and
rotation has not been presented in the literature.
Here we report on such a study. 
We investigate a simplified geometrical configuration corresponding to a
spherical gas cloud with symmetrical radial outflows and solid body
rotation. 
We base our modeling on a Monte-Carlo radiative transfer code.
Besides modeling the impact of joint rotation and outflows, we also
check to what extent the semi-analytical model presented by
\cite{Garavito14} an also be applied in our case.

>>>>

Recent advances in instrumentation have revealed the presence of gas
rotation on vastly different physical scales.
For instance, spatially resolved spectra on compact dwarf galaxies
have measured clear signs of gas showing pure rotation kinematics
\citep{2015A&A...577A..21C,2017A&A...600A.125C} and the recent mapping of
high redshift circumgalactic regions have also revealed kinematic
evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
Systems with star formation, neutral gas and low dust
contents can produce a \lya emission line \citep{PartridgePeebles}
motivating the observational work to phenomenologically link tracers
of galaxy rotation such as H$\alpha$ to \lya spectra
\citep[e.g.][]{Herenz2016}.

@@

However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies are
more complex than pure rotation.
In many observations the \lya line profile has a single peak
redwards from the line's center, in other cases there is a double peak
but the peak on the red side is stronger
\citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
These features have been explained as the consequence of multiple
\lya photon scatterings through a homogeneous outflowing shell of
neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2015ApJ...812..123G}.

Nevertheless, a study of the combined effects of outflows and
rotation has not been presented in the literature.
Here we report on such a study with the main aim of quantifying the
validity of the semi-analytic approximation presented by
\cite{Garavito14} in the case where outflows are also present. 
We investigate a simplified geometrical configuration corresponding to a
spherical gas cloud with symmetrical radial outflows and solid body
rotation and contrast the semi-analytic model against the results of a
Monte-Carlo radiative transfer code.
